---
title: "Analysis Template"
author: "Your Name"
date: last-modified
bibliography: ../source/references.bib
csl: ../source/apa.csl
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
execute:
  echo: true
  warning: false
---

{{< include ../_hardcoded_values.qmd >}}

Download this template from the [GitHub repository][repo-url] or via the `</>Code`-button in the top right corner. This template is designed to be a starting point for your analyses, following many of the best practices for reproducibility and clarity. Note that you may need to download and place additional content in the respective folder for this script to run (e.g. `references.bib`).

# Dependencies

We load all necessary libraries at the beginning to ensure portability and clarity.

::: box-r
```{r}
#| label: setup-r

library(here)    # For relative paths
library(ggplot2) # For visualization
```
:::

::: box-python
```{python}
#| label: setup-python

import os
import pandas as pd
import matplotlib.pyplot as plt
import random
import session_info
```
:::

# Hardcoded Values

We define constants here to avoid "magic numbers" later in the script.

::: box-r
```{r}
#| label: hardcoded-values-r

ATTENTION_THRESHOLD <- 2
PLOT_COLOUR <- "steelblue"
```
:::

::: box-python
```{python}
#| label: hardcoded-values-python

ATTENTION_THRESHOLD = 2
PLOT_COLOUR = "steelblue"
```
:::

# Data Management

## Loading Data

Using relative paths ensures that the script runs on any device.

::: box-r
```{r}
#| label: load-data-r

# Example for a real file: 
my_data <- read.csv(here("data", "raw", "my_data.csv"))
```
:::

::: box-python
```{python}
#| label: load-data-python

# In Python, we go one level up to reach the project root from 'scripts/'
path = os.path.join("..", "data", "raw", "my_data.csv")
df = pd.read_csv(path)
```
:::

## Processing

We follow naming conventions (snake_case) and explain *why* we filter data.

::: box-r
```{r}
#| label: processing-r

# Removing cases that don't meet the attention threshold
cleaned_data <- subset(my_data, reaction_time < ATTENTION_THRESHOLD)
```
:::

::: box-python
```{python}
#| label: processing-python

# Removing cases that don't meet the attention threshold
df_cleaned = df[df['reaction_time'] < ATTENTION_THRESHOLD].copy()
```
:::

We can also save the cleaned data for future use.

::: box-r
```{r}
#| label: save-cleaned-data-r

write.csv(cleaned_data, here("data", "processed", "cleaned_data.csv"))
```
:::

::: box-python
```{python}
#| label: save-cleaned-data-python

save_path = os.path.join("..", "data", "processed", "cleaned_data.csv")
df_cleaned.to_csv(save_path, index=False)
```
:::

# Analysis & Visualization

## Statistical Analysis

Here we would perform our statistical tests, ensuring to report all relevant details for reproducibility.

::: box-r
```{r}
#| label: stats-r

mean_reaction_time <- mean(cleaned_data$reaction_time)
```
:::

::: box-python
```{python}
#| label: stats-python

mean_reaction_time = df_cleaned['reaction_time'].mean()
```
:::

To report the result I can simply use inline code. The mean reaction time is `r round(mean_reaction_time, 2)` seconds.

## Plotting

According to our workflow, we create all figures within the code and save all figures into the `outputs/figs/` directory.

::: box-r
```{r}
#| label: plot-r

# Create the plot
p <- ggplot(cleaned_data, aes(x = condition, y = reaction_time)) +
  geom_point(color = PLOT_COLOUR) +
  theme_minimal() +
  labs(title = "Reaction time in dependence of condition",
       caption = "Data source: Reaction test XY")

# Display
p

# Save using 'here' to target the project root
ggsave(here("outputs", "figs", "analysis_plot_r.png"), plot = p, width = 6, height = 4)
```
:::

::: box-python
```{python}
#| label: plot-python

# Create the plot
plt.figure(figsize=(6, 4))
plt.scatter(df['condition'], df['reaction_time'], color=PLOT_COLOUR, alpha=0.7)
plt.title('Reaction time in dependence of condition')
plt.xlabel('Condition')
plt.ylabel('Reaction Time')

# Display
plt.show()

# Save using os.path to target the project root
save_path = os.path.join("..", "outputs", "figs", "analysis_plot_py.png")
plt.savefig(save_path, dpi=300, bbox_inches='tight')
```
:::

# Reproducibility Extras

## Seeds & Randomness

To ensure results remain the same across devices, we set a seed.

::: box-r
```{r}
#| label: seed-r

set.seed(123)
sample_values <- sample(1:100, 3)
sample_values
```
:::

::: box-python
```{python}
#| label: seed-python

random.seed(123)
sample_values = random.sample(range(1, 101), 3)
sample_values
```
:::

## Citations

As discussed by @wickham2023...

## Equations

$$y = \beta_0 + \beta_1 x + \epsilon$$

## References

::: {#refs}
:::

## Session Information

This section is vital for long-term maintenance and troubleshooting.

::: box-r
```{r}
#| label: info-r

sessionInfo()
```
:::

::: box-python
```{python}
#| label: info-python

session_info.show()
```
:::
