---
title: "Style"
---

{{< include ../_hardcoded_values.qmd >}}


```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false

suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(ggplot2))
```

```{python}
#| eval: true
#| echo: false
#| message: false
#| warning: false

import os
import pandas as pd
import random
from datetime import date
import session_info
import matplotlib.pyplot as plt
```

::: callout-note
To better understand the source code behind the code style conventions it is recommended to click on the `</>Code`-button in the top right corner of this page and select "Show source code". This will show you the exact source code of this page.
:::

# Code style

A **codestyle** is a set of conventions and rules for building script and writing the code within it. Adhering to the following rules is partly mandatory for it to run successfully, to make your scripts easier to read for others (and for your "future self") and ensures that your analysis can be maintained over time.

## Meta settings

Within a script, important meta-information can be provided using the YAML header. This includes not only the author, the date, or the integration of citations, but also formatting options for the rendered document. Here is examplary content of a YAML header:

```{r}
#| eval: false
#| echo: true

---
title: "Analysis of Psychological Well-being"
author: "Your Name"
date: last-modified

bibliography: references.bib # Path to your bibliography file (see section "Citation" in the workflow chapter)
csl: apa.csl # Path to your citation style file (see section "Citation" in the workflow chapter)

format:
  html:
    toc: true              # Adds a Table of Contents
    number-sections: true  # Numbers your headings
    code-fold: true        # Allows readers to click to show/hide code
execute:
  echo: true               # Shows the code in the output by default (global setting)
  warning: false           # Hides R/Python warnings in the final report (global setting)
---
```

## Dependencies

::: callout-note
Make sure to not include the installation of the package (e.g., `install.packages("ggplot2")`) in the script, as this would repeatedly install the packages each time the script is rendered. Instead install packages if necessary directly via the Console.
:::

Always load your libraries/packages at the very top of your script. This acts as a "list of ingredients" so others know what needs to be installed before running your code. Make sure you do not include any unnecessary packages, as this can lead to confusion and potential issues with package conflicts.

The loading of packages can take place within the script:

::: box-r
```{r}
#| eval: true
#| echo: true

library(ggplot2)
```
:::

Python and R do not share the same packages. For data visualization, for example, a commonly used package in Python is `matplotlib`. To load it, you can use the following code:

::: box-python
```{python}
#| eval: true
#| echo: true
#| 
import matplotlib.pyplot as plt
```
:::

## Chunk options

In the IDE of RStudio, settings how chunks should behave globaly can be set within the YAML header. However, these settings can be overwritten for each chunk individually by adding the following options at the beginning of a chunk:

```text
#| eval: true
#| echo: true
#| output: true
#| warning: true
#| error: false
#| include: false
```

The settings of a given chunk can also be set directly within the curly brackets at the beginning of a chunk, for example:

```text
{r, eval = TRUE, echo = TRUE, output = TRUE, warning = TRUE, error = FALSE, include = FALSE}
```

Here are the explanations of the different options:

- **`eval: true`**: The chunk is actually executed/calculated.
- **`echo: true`**: The code of the chunk is visible in the rendered document.
- **`output: true`**: The results of the chunk (e.g., a table) are shown.
- **`warning: true`**: R warnings appear in the document.
- **`error: false`**: If the code within the chunk has an error, rendering stops immediately (safe mode).
- **`include: false`**: The chunk, including code, output, and potential warnings or errors, is completely ignored.

Please note that this is not an extensive list of all available chunk options. For more, see for example the [quarto documentation about code cells and knitr chunk options](https://quarto.org/docs/reference/cells/cells-knitr.html).

## Hardcoded values

Hardcoding refers to the practice of embedding specific data—like numbers, file paths, or fixed thresholds—directly into your analysis code instead of using a variable.

Avoid hardcoded values within your code. Instead, define them as variables within a separate chunk at the beginning of your script. This makes it easier to update values in one place and ensures consistency throughout your analysis. To make these values stand out, use a specific naming convention (e.g., UPPERCASE_LETTERS with underscores).

::: box-r
::: {.panel-tabset}

## GOOD

The value is defined at the beginning of the script in a separate chunk for hardcoded values...

```{r}
#| eval: true
#| echo: true

MAX_SCORE <- 10
```


... and later on refered to in the analysis.
```{r}
#| eval: true
#| echo: true

filtered_seq <- seq(1:MAX_SCORE)
```

## BAD

The value is simply inserted in the middle of your script. If they change you have to search for all of these and change them manually.

```{r}
#| eval: true
#| echo: true

filtered_seq <- seq(1:10)
```
:::
:::

## Visual structure

Headings are the basis for the table of content (toc) in the rendered html-file when such is enabled in the YAML header. They also help to structure the code and make it easier to navigate for others (and for your "future self"). It is recommended to use at least three levels of headings (e.g., `#`, `##`, and `###`) to create a clear structure.

## Portability

A reproducible script should work regardless of which computer it is running on. Absolute paths (e.g., `C:/Users/Username/Documents/...`) are the "enemy" of reproducibility because they only exist on your specific device.

Instead use relative paths which point to a location relative to your project's root folder.

::: box-r
::: {.panel-tabset}

## GOOD

Relative path, even better to use with the function `here()` from the same named package as this ensures the relative path starts at the project file:

```{r}
#| eval: true
#| echo: true

my_data <- read.csv(here("data", "raw", "my_data.csv"))
```

## BAD

Absolute path:

```{r}
#| eval: false
#| echo: true

my_data <- read.csv("C:/Users/Username/Documents/Project/data/raw/my_data.csv")
```
:::
:::


::: box-python
In Python, you can achieve the same using the `os` module to construct relative paths:

Relative path (note that in Python the relative path starts at the scripts directory)

```{python}
#| eval: true
#| echo: true

path = os.path.join("..", "data", "raw", "my_data.csv")
my_data = pd.read_csv(path)
```
:::

## Naming Conventions

Clear naming is the first step toward "self-documenting code."

- **Variables**: Use nouns (e.g., `participant_scores`).

- **Functions**: Use verbs (e.g., `calculate_mean`).

- **Style**: Use snake_case (lowercase letters with underscores) for both R and Python.

Avoid generic names like `data1`, `temp`, or `foo`. Instead, be descriptive: `raw_survey_data` is much more helpful than just `df`.

## Comments

Comments should explain why you are doing something, rather than describing the code itself. Anyone can look up what e.g. the `mean()` function does, but only you know why you chose to exclude certain outliers before calculating it.

::: box-r
::: {.panel-tabset}

## GOOD

Explaining the context of the code, not the logic itself:

```{r}
#| eval: false
#| echo: true

# Removing participants who failed the attention check
cleaned_data <- subset(my_data, attention_check == "passed")
```

## BAD

Explaining basic code logic:

```{r}
#| eval: false
#| echo: true

# Defining a variable and using the subset function
cleaned_data <- subset(my_data, attention_check == "passed") # Result is saved
```
:::
:::

## Seed

If your analysis involves random processes (like creating random groups or running a bootstrap), you **must** set a seed. This ensures that the "random" numbers are the same every time the script is run. this ensures reproducibility across devices and over time.

::: box-r
```{r}
#| eval: true
#| echo: false

set.seed(123)

sample(1:10, 5)
```
:::

::: box-python
```{python}
#| eval: true
#| echo: false

random.seed(123)

random.sample(range(1, 10), 5)
```
:::

## Functions

**Don`t repeat yourself (DRY)**: If you find yourself writing the same code multiple times, it's a sign that you should create a function. This reduces errors—if you need to change a parameter, you only have to change it in one place.

::: box-r
```{r}
#| eval: true
#| echo: false

#' Convert Celsius to Fahrenheit
#' 
#' The function converts a temperature value from Celsius to Fahrenheit.
#' @param celsius A numeric value representing degrees Celsius.
#' @return A numeric value representing degrees Fahrenheit.
# Define the conversion function
convert_celsius_to_fahrenheit <- function(celsius) {
  # Calculate fahrenheit using the standard formula
  fahrenheit <- (celsius * 9/5) + 32
  
  # Return the final result
  return(fahrenheit)
}

# Apply the function to different values
convert_celsius_to_fahrenheit(0)   # Freezing point
convert_celsius_to_fahrenheit(20)  # Room temperature
convert_celsius_to_fahrenheit(37)  # Body temperature
```
:::

::: box-python
```{python}
#| eval: true
#| echo: false

# Define the conversion function
def convert_celsius_to_fahrenheit(celsius):
    """
The function convert_celsius_to_fahrenheit converts a temperature value from Celsius to Fahrenheit using the standard formula F = (C × 9/5) + 32.
Arguments: @param celsius: A numeric value (or vector/series) representing the temperature in degrees Celsius.
Return value: A numeric value (or vector/series) representing the corresponding temperature in degrees Fahrenheit
    """
    # Calculate fahrenheit using the standard formula
    fahrenheit = (celsius * 9/5) + 32
    
    # Return the final result
    return fahrenheit

# Apply the function to different values
print(convert_celsius_to_fahrenheit(0))   # Freezing point
print(convert_celsius_to_fahrenheit(20))  # Room temperature
print(convert_celsius_to_fahrenheit(37))  # Body temperature
```
:::

## Parallelization

Normally, your code runs **sequentially** (one task after another). If you have a computer with multiple CPU cores, **parallelization** allows you to run multiple independent tasks at the same time, significantly reducing execution time for "embarrassingly parallel" tasks.

You can use it, for example, in R (Simulations) if you need to run a lot of independent iterations of the same model or in Python if you want to make many requests to APIs or networks in general.

Consult your supervisor for more information on how to implement parallelization in your specific project, as it can be complex and may not be necessary for all analyses.

## Plots and figures

Plots and figures of your final thesis should never be edited manually but always within the script only. This ensures they are fully reproducible. You can save your plots and figures directly from your script using the `ggsave()` function in R or the `savefig()` function in Python. 

::: box-r
```{r}
#| eval: true
#| echo: true

# Create a simple plot
plot <- ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point() +
  theme_minimal()

# Show plot
plot
```

```{r}
#| eval: true
#| echo: false

# Save the plot
ggsave(here("outputs", "figs", "mtcars_plot.png"), plot = plot, width = 6, height = 4)
```
:::

::: box-python
```{python}
#| eval: true
#| echo: true
 
# Create a simple plot
plt.plot([1, 2, 3], [10, 20, 15], marker='o')

# Show plot
plt.show()
```

```{python}
#| eval: true
#| echo: false

# Save the plot (Python`s root is in the respective file directory, not the project directory)
plt.savefig(os.path.join("..", "outputs", "figs", "own_example_plot.png"))
```
:::

## Equations

If you want to write mathematical formulas or just simple equations in your report (e.g., to explain a statistical model), you can use LaTeX syntax directly within your Quarto document: `$y = \beta_0 + \beta_1 x + \epsilon$.`

For standalone formulas on a new line, use double dollar signs:

`$$y = \beta_0 + \beta_1 x + \epsilon$$`

When rendered, this will display the formula in a professional mathematical format like $y = \beta_0 + \beta_1 x + \epsilon$.

## In-line Code

In-line code allows you to include the results of code directly within your text. This is particularly useful for reporting specific values, such as means, standard deviations, or p-values, without having to create a separate code chunk for it. For R, you can use the ```text `r ` ``` syntax to include in-line code.

This document was last rendered on the ```text r Sys.Date()``` will display as `r Sys.Date()`.

::: box-python
As equivalent for python based in-line code you can simply use ```text `python ` ```.

This document was last rendered on the ``` `python date.today()` ``` will display as `python date.today()`
:::

## Session

At the end of your report, always include a summary of your system and software versions. This is vital for troubleshooting if a package update or software update breaks your code in the future.

::: box-r
```{r}
#| eval: true
#| echo: true

sessionInfo()
```
:::

::: box-python
```{python}
#| eval: true
#| echo: true

session_info.show()
```
:::

::: callout-note
The IDE RStudio offers an automatic formatter. Use `Ctrl + Shift + A` (Windows) or `Cmd + Shift + A` (Mac) to reformat a highlighted code block.

You can also use the `styler` package in R to automatically format your code according to a consistent style.
:::
