---
title: "Workflow"
---

{{< include ../_hardcoded_values.qmd >}}

# Fundamentals

The core idea of a robust, reproducible analysis is a clean, repeatable, and **script-based workflow**. This means that every step of the data analysis is documented in a script and can be executed with a single command. This makes it possible to reproduce the analysis at any time and verify that the results are consistent. The following sections delve deeper into the individual aspects of the script-based workflow:

``` text
Parent_Folder/
├── project.Rproj         # Project file
├── data/                 # All analysis inputs
│   └── raw/              # Raw data
├── scripts/              # All analysis scripts
│   └── analyses.qmd      # Main script
├── outputs/              # All analysis outputs
│   ├── figs/             # Graphics and plots
│   └── tables/           # Tables
├── README.md             # Brief project explanation
├── references.bib        # BibTex file
└── licence.txt           # License information
```

## Project

An RStudio Project (file extension *.Rproj*) is the foundation for a script-based workflow. It can be thought of as the coordination center for the entire project. All files belonging to the same project should be located in the same folder or one of its subfolders. Data can be loaded and saved relative to the project file. This makes it possible to share the entire project folder and verify its reproducibility elsewhere.

To simplify the use of relative paths for loading and saving files, the function `here::here()` can be used.

```{r, eval = FALSE}
# install.packages("here")

here::here("data", "raw", "my_raw_data.csv") # Example path
```

::: callout-note
Project files are also necessary to enable version control via Git and GitHub. You will find more information on this in the following pages.
:::

## Raw data

Raw data is essential for a reproducible and script-based workflow. Raw data refers to materials that could not be recovered if lost (for example, completed questionnaires). Therefore, the first step in a script-based workflow is invariably to create a copy of the raw data, store it under data/raw, and work exclusively with these copies. This ensures that an unaltered version of the raw data can always be accessed if necessary.

## Scripts

As the name of the script-based workflow suggests, the script is the heart of such a workflow. All steps of the data analysis—from data cleaning and transformation to visualization—should be documented in a script. This automatically prevents manual edits and ensures that traceable and reproducible results are obtained. The script connects the (copied) raw data from data/raw and creates the corresponding plots and tables for outputs.

::: callout-note
Citations can also be integrated into the script. Furthermore, version control can be managed directly from within the script. You can find more information about this below under "Citations" and on the following pages. Information regarding what to keep in mind within the scripts can also be found on one of the following pages.
:::

## Outputs

The resulting data should be stored separately from the raw data to avoid confusion and to allow for a fresh start with unaltered raw data if necessary. Outputs may include, for example, plots or tables. It is important that scripts are designed to generate these outputs automatically when executed. This ensures that manual edits are prevented and that traceable, reproducible plots and tables are obtained.

## README

As the name of the README file (file extension *.md*) suggests, this file is intended to be read first when an outside person opens the project. You can think of a README as a manual for the project, which allows third parties to understand the work and build upon it. A README typically includes the following sections:

- *Project Name & Brief Description*: A short overview describing the scope and purpose of the project.

- *Project Content*: A brief description of what is contained in the various folders and how the files are structured.

- *Instructions for Reproduction*: This describes the steps necessary to replicate the results. Reproducibility is often measured by how comprehensive these instructions are and how much prior knowledge a third party requires.

- *Licensing*: If there is no separate license file, the project's licensing terms can be described here.

## Citation

When working within a project file, it is possible to integrate citations directly into the script. All you need is a BibTeX file (file extension *.bib*) containing the corresponding bibliographic entries. This can be created using reference management software like [Zotero](https://www.zotero.org/) or manually.

@book{Grinschgl et al. 2026, title={R you Ready}, author={Grinschgl, Sandra, Hirt, Laura and Schilling, Lars}, year={2026}, publisher={R you Ready Publishing} }

If you want your citations formatted specifically in APA 7th edition style, you must first download the corresponding [Citation Style Language (CSL)](https://www.zotero.org/styles) file and save it in your project folder.

The integration of the BibTeX and CSL files is handled within the YAML header of your Quarto script:

```{r, eval = FALSE}
---
title: "My analysis"
bibliography: references.bib
csl: apa.csl
---
```

In your script, you can then include the citation either directly in the text using `@Grinschgl2026` or in square brackets at the end of a sentence.

## License

Licensing of code and data is an important step to ensure the reusability and protection of one's own work. There are [various types of licenses](https://choosealicense.com/) that regulate whether and in what form one's own work may be reproduced and which requirements must be observed. Depending on the license, a suggestion can also be included on how one's own work should be cited.
